---
title: "Sta602 Homework 5"
author: "Dingyan Zhong"
date: "2023-02-08"
output: pdf_document
---
## Question 1
### (a)
Since Y follows a binomial distribution,
$$
\begin{aligned}
  &p(Y=y|\theta) = \binom{n}{y}\theta^y(1-\theta)^{n-y}\\
  \Rightarrow &\ln(p(Y=y|\theta)) = \ln\binom{n}{y}+y\ln\theta+(n-y)\ln(1-\theta)\\
\end{aligned}
$$
Therefore,
$$
\frac{\partial \ln p(Y=y|\theta)}{\partial \theta} = \frac{y}{\theta}+\frac{n-y}{\theta-1}
$$

$$
\frac{\partial^2 \ln p(Y=y|\theta)}{\partial \theta^2} = -\frac{y}{\theta^2}-\frac{n-y}{(\theta-1)^2}
$$
Then, we have
$$
\begin{aligned}
  I(\theta) &= \sum_{y=0}^n (\frac{y}{\theta^2}+\frac{n-y}{(\theta-1)^2})p(Y=y|\theta)\\
            &= \left(\frac{1}{\theta^2}-\frac{1}{(\theta-1)^2}\right)\sum_{y=0}^n yp(Y=y|\theta)
               +\frac{n}{(\theta-1)^2}\sum_{i=0}^np(Y=y|\theta)\\
            &= \left(\frac{1}{\theta^2}-\frac{1}{(\theta-1)^2}\right)n\theta + \frac{n}{(\theta-1)^2}\\
            &= \frac{n}{\theta(1-\theta)}
\end{aligned}
$$

Eventually,
$$
\begin{aligned}
    p_{J}(\theta) &\propto \sqrt{\frac{n}{\theta(1-\theta)}}\\
                  &\propto \theta^{-\frac{1}{2}}(1-\theta)^{-\frac{1}{2}}\\
                  &\propto Beta(\frac{1}{2},\frac{1}{2})\\
\end{aligned}
$$

### (b)
Similar to part(a),
$$
  \ln p(Y=y|\psi) = \ln \binom{n}{y}+\psi y-n\ln(1+e^{\psi})
$$
Therefore,
$$
\begin{aligned}
  \frac{\partial \ln p(Y=y|\psi)}{\partial \psi} &= y -n\left(1-\frac{1}{1+e^{\psi}}\right)\\
  \frac{\partial^2 \ln p(Y=y|\psi)}{\partial \psi^2} &= -\frac{ne^{\psi}}{(1+e^{\psi})^2}\\
\end{aligned}
$$
Hence,
$$
I(\theta) = \sum_{y=0}^n\frac{ne^{\psi}}{(1+e^{\psi})^2}p(Y=y|\psi) = \frac{ne^{\psi}}{(1+e^{\psi})^2}
$$
and

$$
p_J(\psi) \propto \sqrt{\frac{ne^{\psi}}{(1+e^{\psi})^2}} \propto \frac{e^{\psi/2}}{1+e^{\psi}}
$$

### (c)
From the problem description, we know
$$
  \theta = \frac{e^\psi}{1+e^\psi}
$$
By Change of Variable Theorem,
$$
\begin{aligned}
p_J(\psi) &\propto \left(\frac{e^\psi}{1+e^\psi}\right)^{-1/2}\left(\frac{1}{1+e^\psi}\right)^{-1/2}\left|\frac{d\theta}{d \psi}\right|\\
          &\propto \left(\frac{e^\psi}{1+e^\psi}\right)^{-1/2}\left(\frac{1}{1+e^\psi}\right)^{-1/2}\frac{e^\psi}{(1+e^\psi)^2}\\
          &\propto \frac{e^{\psi/2}}{1+e^{\psi}}
\end{aligned}
$$

## Question 2
### (a)
```{r}
# The known part of posterior
post = function(x) sin(pi*x)^2

# Function for rejection sampling
simulate = function(f = post, envelop = "unif", param1 = 0, param2 = 1, 
                    size = 10^2, has_plot = TRUE){
  r_envelop = match.fun(paste0("r", envelop))
  d_envelop = match.fun(paste0("d", envelop))
  
  # Sample & Reject
  proposal = r_envelop(size, param1, param2)
  density_ratio = post(proposal)/d_envelop(proposal, param1, param2)
  samples = proposal[runif(size)<density_ratio]
  accept_ratio = length(samples)/size
  
  # Draw the histogram for acceptance ratios
  if(has_plot){
    hist(samples, probability = TRUE, xlab = "Sample", ylab = "Density",
         main = paste0("Histogram of ", size, "samples from ", envelop, 
                       "(", param1, ",", param2, "). \n Acceptance ratio: ",
                       round(accept_ratio, 2)), cex.main = 0.75)
  }
  
  return(accept_ratio)
}

```

```{r}
# Visualize results for Unif(0,1) and Beta(2,2)
par(mfrow = c(2,2), mar = rep(4,4))

unif_small_sample = simulate()
unif_large_sample = simulate(size = 10^5)

beta_small_sample = simulate(envelop = "beta", 
                             param1 = 2, param2 = 2)

beta_large_sample = simulate(envelop = "beta", 
                             param1 = 2, param2 = 2, size = 10^5)
```

### (b)
(i)Those methods are almost the same by merely looking at their acceptance rates. But Beta(2,2) seems to be slightly better. The shape of the sample distribution obtained from Beta, when the sample size is small, is more similar to that of the posterior than Uniform.

(ii)One possible reason is that Beta(2,2) has a similar shape as the target posterior. And empirically, setting Beta(2,2) as the envelop function gives us slightly larger acceptance ratios, as is shown below.

```{r}
trial = 10000
accept_unif = rep(0,trial)
accept_beta = rep(0,trial)

for (i in 1:trial){
  accept_beta[i] = simulate(envelop = "beta", param1 = 2, 
                            param2 = 2, has_plot = FALSE)
  accept_unif[i] = simulate(has_plot = FALSE)
}

par(mfrow = c(1,2))
hist(accept_beta, xlab = "Acceptance_Ratio", ylab = "Density", 
     main = "Distribution of Acceptance Ratios: \n Beta")
hist(accept_unif, xlab = "Acceptance_Ratio", ylab = "Density", 
     main = "Distribution of Acceptance Ratios: \n Uniform")
```

